```{r}
options(repos = c(CRAN = "https://cloud.r-project.org")) # set CRAN mirror
```

---
title: "Take Home Exercise 3"
subtitle: "Test a (causal) hypothesis with observational data (THE III)"
author: "Nicolas Waser"
date: "2024-12-12"
format:
  pdf:
    toc: false
    number-sections: true
    number-depth: 3
    colorlinks: true
editor: 
  markdown: 
    wrap: 72
---

*Disclaimer! For this exercise I made use of the following AI Generation
tools: Deepl.com for translation purposes, the integrated Copilot tool
in R Studio for generating code and text and the premium version of
OpenAI's GPT-4o, as well as the free versions of Anthropics Claude 3.5
Sonnet and Mistral's Codestral AI model for generating code and
assisting me in solving the tasks.*

***Exercise 1:** Draw a DAG of the causal hypothesis You will use the
CSES IMD data to test one of the most prominent theories in political
science:the economic voting theory. This theory states that citizens
decide their vote based on the state of the economy, so that they are
more likely to support the incumbent (i.e., the outgoing governing party
or candidate) if the economy works well and less likely is the economic
situation is bad (for a review, see Lewis-Beck and Stegmaier, 2018).
Focusing on citizens’ perceptions rather than objective economic
indicators, the following (causal) hypothesis follows:*

***Hypothesis:** Negative (positive) economic evaluations reduce
(increase) the probability of voting for the incumbent.*

*The CSES data contains one variable that we can use as our independent
(“treatment”) variable, as it is operationalized as follows:*

*• **IMD3013_1:** Would you say that over the past twelve months, the
state of the economy in \[COUNTRY\] has gotten better, stayed about the
same, or gotten worse?*

*It also contains a variable that we can use to measure voting for the
incumbent (i.e., our dependent variable):*

*• **IMD3002_OUTGOV:** Whether or not the respondent cast a ballot for
the outgoing incumbent.*

*Since we want to approximate a causal test, we will need to control for
some variables in order to rule out endogeneity concerns. However, we do
not want to do this blindly but based on theoretical reasoning. As you
will see in the codebook, there are many variables that we can use (and
more that we can construct) to try to isolate the causal relationship
between economic evaluations and voting for the incumbent, but we do not
want to use them all. To guide our model specification, **please draw a
DAG of the theoretical relationship at hand.** This will guide our next
choices, so it is important that it is done carefully.*

*You can draw a DAG in R using the packages dagitty and ggdag (see
installation and loading below). Alternatively, you can draw it with any
other program or by hand and upload it to the .qmd document as an image.
Please comment your decisions.*

*PS: Remember to simplify and do not include every variable you think it
could be involved in the relationship, but only the most important.
Also, group variables under broader concepts to avoid overfitting the
DAG (e.g., ‘socio-economic conditions’ instead of ‘employment status’
and ‘income’).*

```{r}
library(tidyverse)
install.packages("dagitty")
library(dagitty)
install.packages("ggdag")
library(ggdag)
```

***Exercise 2:** Prepare your data*

*Based on your DAG, declare what variables you will use for the
analysis. Please inspect and transform them when necessary (e.g., recode
missing values, inverse scales, etc.). Report your code and comment your
decisions.*

***Exercise 3:** Test your hypothesis with an ordinary-least squares
(OLS) multiple regression model*

*Run an OLS regression model to test your hypothesis. Use the function
lm() for that, or you can use more complicated functions if preferred.
Then comment on your choices and your results. Based on them, does the
evidence support your hypothesis?*

*Optional: you can plot the predicted probabilities of voting for the
incumbent based on economic evaluations. This may be helpful for
interpreting your results.*

#####AI

**Interpretation of the Linear Probability Model Results**

In this model, the dependent variable (Vote_For_Incu) is likely a binary
indicator of whether the individual voted for the incumbent. Because
this is a linear probability model (LPM), the coefficients can be
interpreted as changes in the probability of voting for the incumbent
associated with a one-unit change in the explanatory variables, holding
other variables constant.

	1.	**Eco_Eval (Economic Evaluation)** **Coefficient**: -0.0796

This suggests that a one-unit increase in the respondent’s economic
evaluation measure is associated with about a 7.96 percentage point
decrease in the probability of voting for the incumbent, all else equal.

The strong negative and highly statistically significant coefficient (p
\< 0.001) indicates that as economic evaluation increases (depending on
how it’s coded—if higher values mean a more negative or positive
assessment of the economy, it’s crucial to be clear on the scaling), the
likelihood of supporting the incumbent decreases.

	2.	**Employment** **Coefficient**: -0.0023

This coefficient is not statistically significant (p \> 0.05), meaning
changes in the Employment variable do not reliably predict changes in
the probability of voting for the incumbent. Effectively, employment, as
measured here, doesn’t have a clear linear relationship with incumbent
support in this sample.

	3.	**Partisan (Partisanship)** **Coefficient**: 0.0225

This indicates that a one-unit increase in the partisan measure is
associated with a 2.25 percentage point increase in the probability of
voting for the incumbent, all else equal. The positive and statistically
significant effect (p \< 0.001) suggests that stronger partisanship
(likely favoring the incumbent’s party) boosts the likelihood of
supporting that incumbent.

	4.	**Sat_with_Dem (Satisfaction with Democracy)** **Coefficient**:
-0.0438

A one-unit increase in satisfaction with democracy is associated with a
4.38 percentage point decrease in the probability of voting for the
incumbent, holding other factors constant. This effect is strongly
statistically significant (p \< 0.001). If higher values mean greater
satisfaction, this result might be counterintuitive and could suggest
the need to double-check the direction of coding. For example, it might
be that those who are more satisfied with how democracy is functioning
do not feel a need to vote for the incumbent, or there’s some underlying
coding or contextual factor at play. Re-examining how Sat_with_Dem was
coded and interpreted is advisable.

**Model Fit**: **R-squared**: About 0.0939

This suggests that the model explains around 9.39% of the variation in
the outcome. For an individual-level binary outcome, this may not be
unusual, but it indicates that there are many other factors influencing
the decision to vote for the incumbent not captured here.

**F-statistic and p-value**:

The model’s overall significance is very high (p \< 0.001), meaning that
the model as a whole is statistically different from a model with no
predictors. However, the relatively low R-squared shows limited
explanatory power.

**Conclusion**:

Economic evaluation and satisfaction with democracy are strongly and
negatively related to the probability of voting for the incumbent, which
is somewhat surprising and should prompt a closer examination of
variable coding and measurement.

Partisanship has a clear positive effect: those aligned more strongly
with the incumbent’s party are more likely to vote for that incumbent.

Employment is not a significant predictor in this model.

\#####

***Exercise 4 (additional):** Run an additional analysis of your choice*

*This exercise is not mandatory, but it serves only to opt for the
maximum grade (6).*

*Is there any other statistical test you could run to further support or
disprove the hypothesis?Please think on the observable implications of
the theory that could be tested with the CSESIMD data and provide an
additional test. It can be either another regression specification or a
different statistical analysis. Finally, comment on your decisions and
results, and discuss them together with the results of the previous
exercise. You can be as creative as you want here; it is the final
exercise, so enjoy yourself!*
